---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

[![Travis-CI Build Status](https://travis-ci.org/adamhsparks/getCRUCL2.0.svg?branch=master)](https://travis-ci.org/)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/adamhsparks/getCRUCL2.0?branch=master&svg=true)](https://ci.appveyor.com/project/adamhsparks/getCRUCL2.0)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/getCRUCL2.0)](https://cran.r-project.org/package=getCRUCL2.0)

# getCRUCL2.0
## Download and Use CRU CL2.0 Climatology Data in R

Author/Maintainer: Adam Sparks

The getCRUCL2.0 package provides two functions that automate downloading and
importing CRU CL2.0 climatology data into R, facilitate the calculation of
minimum temperature and maximum temperature, and formats the data into a tidy
data frame or a list of raster stack objects for use in an R session. CRU CL2.0
data are a gridded climatology of 1961-1990 monthly means released in 2002 and
cover all land areas (excluding Antarctica) at 10 arc-second resolution. For
more information see the description of the data provided by the University of
East Anglia Climate Research Unit (CRU),
<http://www.cru.uea.ac.uk/cru/data/hrg/tmc/readme.txt>.

## Note

This package automatically converts elevation values from kilometres to metres.

This package crops all spatial outputs to an extent of ymin = -60, ymax = 85,
xmin = -180, xmax = 180. Note that the original wind data include land area for
parts of Antarctica.
    
License: MIT + file LICENSE

Imports: curl, dplyr, plyr, purrr, raster, tidyr, utils

Quick Start
===========

Install
-------

### Development version

A development version is available from from GitHub. If you wish to
install the development version that may have new features (but also may
not work properly), install the [devtools package](https://CRAN.R-project.org/package=devtools), available from CRAN. I strive to keep the master branch on GitHub functional and
working properly, although this may not always happen.

If you find bugs, please file a [report as an issue](https://github.com/adamhsparks/getCRUCL2.0/issues).

``` r
install.packages("devtools")
devtools::install_github("adamhsparks/getCRUCL2.0", build_vignettes = TRUE)
```

Using getCRUCL2.0
-----------

## Creating tidy data frames for use in R
Create a tidy data frame of all CRU CL2.0 climatology elements available.

```{r, eval=FALSE}
library(getCRUCL2.0)

CRU_data <- create_CRU_df(pre = TRUE,
                          pre_cv = TRUE,
                          rd0 = TRUE,
                          tmp = TRUE,
                          dtr = TRUE,
                          reh = TRUE,
                          tmn = TRUE,
                          tmx = TRUE,
                          sunp = TRUE,
                          frs = TRUE,
                          wnd = TRUE,
                          elv = TRUE)
```

Create a tidy data frame of mean temperature and relative humidity.
```{r, eval=FALSE}
t_rh <- create_CRU_df(tmp = TRUE,
                      reh = TRUE)
```

## Creating raster stacks for use in R
The `create_CRU_stack` function provides functionality for producing a list of
[raster](https://CRAN.R-project.org/package=raster) stack objects for use in an
R session.

Create a list of raster stacks of all CRU CL2.0 climatology elements available.

```{r, eval = FALSE}
CRU_stack <- create_CRU_stack(pre = TRUE,
                              pre_cv = TRUE,
                              rd0 = TRUE,
                              tmp = TRUE,
                              dtr = TRUE,
                              reh = TRUE,
                              tmn = TRUE,
                              tmx = TRUE,
                              sunp = TRUE,
                              frs = TRUE,
                              wnd = TRUE,
                              elv = TRUE)
```

Create a list of raster stacks of maximum and minimum temperature.
```{r, eval=FALSE}
tmn_tmx <- create_CRU_stack(tmn = TRUE,
                            tmx = TRUE)
```

# Reference and abstract

> Mark New (1,*), David Lister (2), Mike Hulme (3), Ian Makin (4)  
A high-resolution data set of surface climate over global land areas
Climate Research, 2000, Vol 21, pg 1-25  
(1) School of Geography and the Environment, University of Oxford, 
    Mansfield Road, Oxford OX1 3TB, United Kingdom  
(2) Climatic Research Unit, and (3) Tyndall Centre for Climate Change Research,
    both at School of Environmental Sciences, University of East Anglia,
    Norwich NR4 7TJ, United Kingdom  
(4) International Water Management Institute, PO Box 2075, Colombo, Sri Lanka

> **ABSTRACT:** We describe the construction of a 10-minute latitude/longitude data
set of mean monthly surface climate over global land areas, excluding Antarctica.
The climatology includes 8 climate elements - precipitation, wet-day frequency,
temperature, diurnal temperature range, relative humidity,sunshine duration,
ground frost frequency and windspeed - and was interpolated from a data set
of station means for the period centred on 1961 to 1990. Precipitation was first
defined in terms of the parameters of the Gamma distribution, enabling the 
calculation of monthly precipitation at any given return period. The data are
compared to an earlier data set at 0.5 degrees latitude/longitude resolution and
show added value over most regions. The data will have many applications in
applied climatology, biogeochemical modelling, hydrology and agricultural
meteorology and are available through the School of Geography Oxford 
(http://www.geog.ox.ac.uk), the International Water Management Institute
"World Water and Climate Atlas" (http://www.iwmi.org) and the Climatic
Research Unit (http://www.cru.uea.ac.uk).

# Notes    
Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md).
By participating in this project you agree to abide by its terms.
