---
title: "getCRUCLdata"
author: "Adam H. Sparks"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{getCRUCLdata}
  %\VignetteEncoding{UTF-8}
---

# Introduction to getCRUCLdata

The getCRUCLdata package provides two functions that automate downloading and
importing CRU CL2.0 climatology data into R, facilitate the calculation of
minimum temperature and maximum temperature, and formats the data into a tidy
data frame or a list of raster stack objects for use in an R session. CRU CL2.0
data are a gridded climatology of 1961-1990 monthly means released in 2002 and
cover all land areas (excluding Antarctica) at 10-minute resolution. For
more information see the description of the data provided by the University of
East Anglia Climate Research Unit (CRU),
<https://crudata.uea.ac.uk/cru/data/hrg/tmc/readme.txt>.

# Using getCRUCLdata
Both functions, `create_CRU_df` and `create_CRU_stack` take the same arguments
but return different types of objects in the R session. Logical arguments are
used to specify the data to retrieve and parse. The arguments for selecting the
climatology elements for download are:  

* **pre** Logical. Fetch precipitation (milimetres/month) from server and
return in the data? Defaults to FALSE.  
* **pre_cv** Logical. Fetch cv of precipitation (percent) from server and
return in the data? Defaults to FALSE.  
* **rd0** Logical. Fetch wet-days (number days with >0.1milimetres rain per
month) and return in the data? Defaults to FALSE.  
* **dtr** Logical. Fetch mean diurnal temperature range (degrees Celsius)
and return it in the data? Defaults to FALSE.  
* **tmp** Logical. Fetch temperature (degrees Celsius) and return it in the
data? Defaults to FALSE.  
* **tmn** Logical. Calculate minimum temperature values (degrees Celsius)
and return it in the data? Defaults to FALSE.  
* **tmx** Logical. Calculate maximum temperature (degrees Celsius) and return
it in the data? Defaults to FALSE.  
* **reh** Logical. Fetch relative humidity and return it in the data?
Defaults to FALSE.  
* **sunp** Logical. Fetch sunshine, percent of maximum possible (percent of
day length) and return it in data? Defaults to FALSE.  
* **frs** Logical. Fetch ground-frost records (number of days with ground-
frost per month) and return it in data? Defaults to FALSE.  
* **wnd** Logical. Fetch 10m wind speed (metres/second) and return it in the
data? Defaults to FALSE.  
* **elv** Logical. Fetch elevation (and convert to metres from kilometres) and
return it in the data? Defaults to FALSE.  

## Creating tidy data frames for use in R

The `create_CRU_df()` function creates tidy data frames of the CRU CL2.0 climatology elements. Illustrated here, create a tidy data frame of all CRU CL2.0 climatology elements available.

```{r, eval=FALSE}
library(getCRUCLdata)

CRU_data <- create_CRU_df(pre = TRUE,
                          pre_cv = TRUE,
                          rd0 = TRUE,
                          tmp = TRUE,
                          dtr = TRUE,
                          reh = TRUE,
                          tmn = TRUE,
                          tmx = TRUE,
                          sunp = TRUE,
                          frs = TRUE,
                          wnd = TRUE,
                          elv = TRUE)
```

Create a tidy data frame of mean temperature and relative humidity.

```{r, eval=FALSE}
t_rh <- create_CRU_df(tmp = TRUE,
                      reh = TRUE)
```

## Creating raster stacks for use in R or saving for use in a GIS
The `create_CRU_stack()` function provides functionality for producing a list of [raster](https://CRAN.R-project.org/package=raster) stack objects for use in an R session.

Create a list of raster stacks of all CRU CL2.0 climatology elements available.

```{r, eval = FALSE}
CRU_stack <- create_CRU_stack(pre = TRUE,
                              pre_cv = TRUE,
                              rd0 = TRUE,
                              tmp = TRUE,
                              dtr = TRUE,
                              reh = TRUE,
                              tmn = TRUE,
                              tmx = TRUE,
                              sunp = TRUE,
                              frs = TRUE,
                              wnd = TRUE,
                              elv = TRUE)
```

Create a list of raster stacks of maximum and minimum temperature.
```{r, eval=FALSE}
tmn_tmx <- create_CRU_stack(tmn = TRUE,
                            tmx = TRUE)
```

Saving raster format files to a local disk as geotiff files in the `tmp` directory with a tmn or tmx prefix to the month for a filename.

```{r, eval=FALSE}
dir.create(file.path("~/tmp"), showWarnings = FALSE)
writeRaster(tmn_tmx$tmn, filename = paste0("~/tmp/tmn_", names(tmn_tmx$tmn)), bylayer = TRUE, format = "GTiff")

writeRaster(tmn_tmx$tmx, filename = paste0("~/tmp/tmx_", names(tmn_tmx$tmn)), bylayer = TRUE, format = "GTiff")
```

# Reference and abstract

> Mark New (1,*), David Lister (2), Mike Hulme (3), Ian Makin (4)  
A high-resolution data set of surface climate over global land areas
Climate Research, 2000, Vol 21, pg 1-25  
(1) School of Geography and the Environment, University of Oxford, 
    Mansfield Road, Oxford OX1 3TB, United Kingdom  
(2) Climatic Research Unit, and (3) Tyndall Centre for Climate Change Research,
    both at School of Environmental Sciences, University of East Anglia,
    Norwich NR4 7TJ, United Kingdom  
(4) International Water Management Institute, PO Box 2075, Colombo, Sri Lanka

> **ABSTRACT:** We describe the construction of a 10-minute latitude/longitude data
set of mean monthly surface climate over global land areas, excluding Antarctica.
The climatology includes 8 climate elements - precipitation, wet-day frequency,
temperature, diurnal temperature range, relative humidity,sunshine duration,
ground frost frequency and windspeed - and was interpolated from a data set
of station means for the period centred on 1961 to 1990. Precipitation was first
defined in terms of the parameters of the Gamma distribution, enabling the 
calculation of monthly precipitation at any given return period. The data are
compared to an earlier data set at 0.5 degrees latitude/longitude resolution and
show added value over most regions. The data will have many applications in
applied climatology, biogeochemical modelling, hydrology and agricultural
meteorology and are available through the School of Geography Oxford 
(http://www.geog.ox.ac.uk), the International Water Management Institute
"World Water and Climate Atlas" (http://www.iwmi.org) and the Climatic
Research Unit (http://www.cru.uea.ac.uk).

